---
title: "In-class Exercise 7"
author: "Chang Fang Yu"
date-modified: "last-modified"
execute:
  echo: True
  eval: True
  warning: false
  freeze: true
---

```{r}
pacman::p_load(tidyverse, tsibble, feasts, fable, seasonal)
```
```{r}
ts_data <- read_csv("visitor_arrivals_by_air.csv")
```
### Conventional base ts object versus tibble object
```{r}
ts_data$'Month-Year' <- dmy(
  ts_data$ 'Month-Year'
)
```
```{r}
ts_data
```
```{r}
ts_data_ts <- ts(ts_data)       
head(ts_data_ts)
```
### tsibble object
```{R}
ts_tsibble <- ts_data %>%
  mutate(Month = yearmonth(`Month-Year`)) %>%
  as_tsibble(index = `Month`)
```

## Visualising Time-series Data
### Visualising single time-series: ggplot2 methods
```{r}
ts_longer <- ts_data %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
ts_longer %>%
  filter(Country == "Taiwan") %>%
  ggplot(aes(x = `Month-Year`, 
             y = Arrivals))+
  geom_line(size = 0.5)
```
### Plotting multiple time-series data with ggplot2 methods

```{r}
#| fig-height: 8
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals,
           color = Country))+
  geom_line(size = 0.5) +
  theme(legend.position = "bottom", 
        legend.box.spacing = unit(0.5, "cm"))

```
```{r eval=FALSE}
#| ecoho: false
#| fig-height: 12
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals))+
  geom_line(size = 0.5) +
  facet_wrap(~ Country,
             ncol = 3,
             scales = "free_y") +
  theme_bw()
```
## Visual Analysis of Time-series Data
```{r}
tsibble_longer <- ts_tsibble %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```
### Visual Analysis of Seasonality with Seasonal Plot
```{r}
#| ecoho: false
#| fig-height: 6
tsibble_longer %>%
  filter(Country == "Italy" |
         Country == "Vietnam" |
         Country == "United Kingdom" |
         Country == "Germany") %>% 
  gg_season(Arrivals)
```
###  Visual Analysis of Seasonality with Cycle Plot

```{R}
#| ecoho: false
#| fig-height: 6
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  autoplot(Arrivals) + 
  facet_grid(Country ~ ., scales = "free_y")
```


```{R}
#| ecoho: false
#| fig-height: 6
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  gg_subseries(Arrivals)
```
## Time series decomposition
### Single time series decomposition
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  ACF(Arrivals) %>% 
  autoplot()
```
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  PACF(Arrivals) %>% 
  autoplot()
```
### Multiple time-series decomposition
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  ACF(Arrivals) %>%
  autoplot()
```
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  PACF(Arrivals) %>%
  autoplot()
```
### Composite plot of time series decomposition
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  gg_tsdisplay(Arrivals)
```

## Visual STL Diagnostics
### Visual STL diagnostics with feasts

```{R}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

### Classical Decomposition with feasts
```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(
    classical_decomposition(
      Arrivals, type = "additive")) %>%
  components() %>%
  autoplot()
```
## Visual Forecasting
### Time Series Data Sampling
```{r}
vietnam_ts <- tsibble_longer %>%
  filter(Country == "Vietnam") %>% 
  mutate(Type = if_else(
    `Month-Year` >= "2019-01-01", 
    "Hold-out", "Training"))
vietnam_train <- vietnam_ts %>%
  filter(`Month-Year` < "2019-01-01")
```


### Exploratory Data Analysis (EDA): Time Series Data
```{R}
vietnam_train %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```
### Fitting forecasting models
```{R}
fit_ses <- vietnam_train %>%
  model(ETS(Arrivals ~ error("A") 
            + trend("N") 
            + season("N")))
fit_ses
```
#### Examine Model Assumptions
```{R}
gg_tsresiduals(fit_ses)
```
#### The model details
```{R}
fit_ses %>%
  report()
```
#### Trend method
```{R}
vietnam_H <- vietnam_train %>%
  model(`Holt's method` = 
          ETS(Arrivals ~ error("A") +
                trend("A") + 
                season("N")))
vietnam_H %>% report()
```
#### Damped Trend Method
```{R}
vietnam_HAd <- vietnam_train %>%
  model(`Holt's method` = 
          ETS(Arrivals ~ error("A") +
                trend("Ad") + 
                season("N")))
vietnam_HAd %>% report()
```
#### Result
```{R}
gg_tsresiduals(vietnam_H)
```
```{R}
gg_tsresiduals(vietnam_HAd)
```